\hypertarget{intro}{%
\section{Intro}\label{intro}}

A complex system is hard to simplify or cannot be simplified at all.

\hypertarget{complicated-vs.complex-system}{%
\subsection{Complicated vs.~complex
system}\label{complicated-vs.complex-system}}

A \textbf{complex system} has a lot of elements and it's difficult to
explain the system, but not because of the elements. The problem is the
correlation between the elements. It's not possible to cut the system
into subsystems due to loops. It would change the behaviour of the
system.\\
\emph{Example: Company}\\
It's not possible to understand a company if you'll only know a part of
it.

If a system has many elements and is therefore not easy to explain, we
have a \textbf{complicated system}. A complicated system is
``reducible'' by cutting it into small systems (subsystems), which can
be explained.\\
\emph{Example: Car}\\
It's a complicated system, but the engine can be analysed seperatly (or
even be splited into more subsystems).

\hypertarget{features-of-a-complex-system}{%
\subsection{Features of a complex
system}\label{features-of-a-complex-system}}

There are no typical features of a complex system, but it often shows
the following properties:

\begin{itemize}
\tightlist
\item
  High number of elements
\item
  Non-linear interactions between the elements
\item
  Delayed effects
\item
  Negative and positive feedback
\item
  Network-like structure
\item
  Open
\item
  Universal
\item
  Dynamic
\item
  Robust
\item
  Creative and innovative
\item
  Unpredictable
\item
  Differentiated sensitivity
\item
  Not monitorable
\end{itemize}

\hypertarget{models-and-simulation}{%
\subsection{Models and simulation}\label{models-and-simulation}}

To manage complexity it is required to build models. A model is a
simplified description of reality.\\
The usefulness of a model is, to test changes within a model and
transfer the gained knowledge to the real world. If this is not
possible, the model is useless.\\
A model must be as accurate as required, but an overengineered model is
also useless. Details must be erased by a model.\\
A model has boundaries it cannot be used for simulations it's not
intended to!

The usage of a model to determine how things are going to act in a
certain situation (in all likelihood), is called \textbf{simulation}.
